\TheTitle

### 2019-02-22 Paper1/framework.tex

\section{Discretization and code framework}
\label{sec:framework}
\subsection{Discretization}
In the following, we discretize the moment system in space and time following according to \cite{kusch2017maximum}. Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:IPMmomentSystem}.  
Omitting initial conditions and assuming a one-dimensional spatial domain, we can write this system  as
\begin{align*}
\partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
\end{align*}
with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition. Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments. First we perform the discretization of the spatial domain: We choose the discrete unknowns to be the spatial averages over each cell at time $t_n$, given by
\begin{align*}
\bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
\end{align*}
If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$ as
\begin{align}\label{eq:IPMDiscretization}
\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
\end{align}
for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\bm{u},\bm{u})=\bm{F}(\bm{u})$.
To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.

When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:origProblem}, then for the IPM system we can simply take
\begin{align*}
 \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle g(u_{ME}(\hat\Lambda(\bm{\hat u}_{j}^{n})),u_{ME}(\hat\Lambda(\bm{\hat u}_{j+1}^{n})))\bm{\varphi}\rangle.
\end{align*}
This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.
The time update of the moment vector now becomes
\begin{align}\label{eq:exactUpdate}
\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}- \frac{\Delta t}{\Delta x}\left( \langle g(u_{ME}(\hat \Lambda_j^n),u_{ME}(\hat \Lambda_{j+1}^n))\bm{\varphi}\rangle - \langle g(u_{ME}(\hat \Lambda_{j-1}^n),u_{ME}(\hat \Lambda_{j}^n))\bm{\varphi} \rangle\right),
\end{align}
where $\hat\Lambda_{j}^n :=\hat\Lambda(\bm{\hat u}_{j}^{n})$ for all $j$. Note that the computation of $\hat\Lambda_{j}^n$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$.

>>>>

\section{Discretization and code framework}
\label{sec:framework}
\subsection{Discretization}
In the following, we discretize the moment system in space and time following according to \cite{kusch2017maximum}. Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system \eqref{eq:IPMmomentSystem}.  
Omitting initial conditions and assuming a one-dimensional spatial domain, we can write this system  as
\begin{align*}
\partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
\end{align*}
with the flux $\bm{F}:\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$, $\bm{F}(\bm{\hat u})=\langle f(\mathcal{U}_{ME}(\bm{\hat u}))\bm{\varphi} \rangle$, where we omit the index ME in the following for efficiency of exposition. Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments. First we perform the discretization of the spatial domain: We choose the discrete unknowns to be the spatial averages over each cell at time $t_n$, given by
\begin{align*}
\bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
\end{align*}
If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{0j}^n,\cdots,\bm{\hat u}_{Nj}^n)^T\in\mathbb{R}^{N+1}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{(N+1)\times p}\times\mathbb{R}^{(N+1)\times p}\to\mathbb{R}^{(N+1)\times p}$ as
\begin{align}\label{eq:IPMDiscretization}
\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
\end{align}
for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$, where $N_x$ is the number of spatial cells and $N_t$ is the number of time steps.
The numerical flux is assumed to be consistent, i.e., that $\bm{G}(\bm{u},\bm{u})=\bm{F}(\bm{u})$.
To ensure stability, a CFL condition has to be derived by investigating the eigenvalues of $\nabla \bm{F}$.

When a consistent numerical flux $\bm g:\mathbb{R}^p\times\mathbb{R}^p\to\mathbb{R}^p$, $\bm g = \bm g(\bm u_\ell, \bm u_r)$ is available for the deterministic problem \eqref{eq:hyperbolicProblem}, then for the IPM system we can simply take
\begin{align*}
 \bm{G}(\bm{\hat u}_{j}^n,\bm{\hat u}_{j+1}^{n}) = \langle \bm g(\mathcal{U}_{ME}(\bm{\hat u}_j^n),\mathcal{U}_{ME}(\bm{\hat u}_{j+1}^n))\bm{\varphi}\rangle.
\end{align*}
This choice of the numerical flux is a common choice in kinetic theory and is called kinetic flux.

The numerical flux requires evaluating the closure $\mathcal{U}_{ME}(\bm{\hat u}_j^n)$. To simplify notation, we define
\begin{align*}
\bm{u}_{ME}(\bm\Lambda):=\left( \nabla_{\bm{\hat u}} s \right)^{-1}(\bm\Lambda),
\end{align*}
meaning that we can write the closure as
\begin{align*}
\mathcal{U}_{ME}(\bm{\hat u}_j^n) = \bm{u}_{ME}(\bm{\hat{\lambda}}(\bm{\hat u}_j^n)^T \bm{\varphi}).
\end{align*}
The computation of the so called entropy variables $\bm{\hat\lambda}_j^n:=\bm{\hat\lambda}(\bm{\hat u}_j^n)$ requires solving the dual problem \eqref{eq:dualProblem} for the moment vector $\bm{\hat u}_{j}^{n}$. Commonly, this optimization problem is solved with Newton's method:

### 2019-02-22 Paper1/main.tex
continued

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

@@

\begin{abstract}
Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques. Intrusive methods require the implementation of a new code and yield

\end{abstract}

\begin{keyword}
conservation laws, hyperbolic, intrusive, UQ, IPM, SG, Collocation
\end{keyword}

@@

\input{introduction}
\input{framework}
\input{colloc_ipm}
\input{one_shot_ipm}
\input{results}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}[theorem]{Remark}

@@

\begin{abstract}
Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques. A main drawbacks of intrusive methods is the need to implement new code, whereas collocation methods recycle a given deterministic solver. Furthermore, intrusive methods come with increased computational costs, making it hard to compete with non-intrusive methods.
In this work, we present a code framework, which facilitates the implementation of intrusive methods by recycling parts of deterministic codes. Additionally, we introduce methods to decrease numerical costs of intrusive methods for steady problems. We demonstrate the effectiveness of the proposed strategies by comparing results of the uncertain NACA0012 testcase as well as the shallow water equations.
\end{abstract}

\begin{keyword}
uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation
\end{keyword}

@@

\input{introduction}
\input{framework}
\input{colloc_ipm}
\input{one_shot_ipm}
\input{results}
\input{summary_outlook}
